---
title: "AQI Datasets"
author: "Brian Jo Hsuan Lee"
date: "11/23/2021"
output: html_document
---

See description for the following variables:

* `days` - Number of days in the year having an Air Quality Index value
* `good` - Number of days in the year having an AQI value 0 through 50
* `moderate` - Number of days in the year having and AQI value 51 through 100
* `unhealthy_for_sensitive_groups` - Number of days in the year having an AQI value 101 through 150
* `unhealthy` - Number of days in the year having an AQI value 151 through 200
* `very unhealthy` - Number of days in the year having an AQI value 201 or higher, which includes the AQI categories very unhealthy and hazardous
* `max_aqi` - The highest daily AQI value in the year
* `90%_aqi` - 90 percent of daily AQI values during the year were less than or equal to the 90th percentile value
* `med_aqi` - Half of daily AQI values during the year were less than or equal to the median value, and half equaled or exceeded it
* `co`, `no2`, `o3`, `so2`, `pm2_5`, `pm10` - A daily index value is calculated for each air pollutant measured. The highest of those index values is the AQI value, and the pollutant responsible for the highest index value is the "Main Pollutant." These columns give the number of days each pollutant measured was the main pollutant.


Load libraries and set working directory
```{r, message=FALSE}
library(tidyverse)
setwd("~/Desktop/Columbia/Fall_2021/P8105-Data_Science/")
```

.csv tidying function 
```{r, warning=FALSE}
tidy_df = function(all_paths){
  read_csv(all_paths, col_types = "fcddddddddddddddd") %>% 
  janitor::clean_names() %>% 
  separate(county, into = c("county", "state"), sep = ", ") %>% 
  rename(
    days = number_days_with_aqi,
    max_aqi = aqi_maximum, 
    `90%_aqi` = aqi_90th_percentile,
    median_aqi = aqi_median,
    co = number_days_co,
    no2 = number_days_no2,
    o3 = number_days_o3,
    so2 = number_days_so2,
    pm2_5 = number_days_pm2_5,
    pm10 = number_days_pm10
  ) %>% 
  mutate(
    county = str_replace(county, " County", ""),
    county = factor(county),
    state = factor(state)
  )
}
```

Data import and organization
```{r, warning=FALSE}
folder_path = "./aqi/"

aqi_files = 
  tibble(
    year = rep(2001:2021, each = 4), 
    st = rep(c("ny", "oh", "me", "pa"), 21),
    path = str_c(folder_path, st, "_aqireport", year, ".csv")
  ) %>% 
  mutate(
    files = map(path, tidy_df)
  ) %>% 
  unnest() %>% 
  select(-(st:county_code)) %>% 
  arrange(year, state, county)

head(aqi_files, 5)
```

