---
title: "Looking at the numbers"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
library(usmap)
library(ggplot2)
library(maps)
library(mapdata)
library(plotly)
library(rjson)
```

## Data Sources

For weather data...

For health outcomes data, we are interested in the incidence of asthma, skin 
cancer and lung cancer across time and space. We obtained age-adjusted incidence 
rates data for these three health outcomes for the U.S. states New York, Ohio 
and Pennsylvania over multiple years. The following bullet points provide links 
from where we obtained the data.

*   Asthma
    *   Provide source
*   Melanoma of the skin (skin cancer)
    *   [New York](https://www.health.ny.gov/statistics/cancer/registry/table2/tb2melanomanys.htm)
    *   [Ohio](https://publicapps.odh.ohio.gov/EDW/DataBrowser/Browse/StateLayoutLockdownCancers)
    *   [Pennsylvania](https://www.phaim1.health.pa.gov/EDD/)
*   Lung cancer
    *   [New York](https://www.health.ny.gov/statistics/cancer/registry/table2/tb2lungnys.htm)
    *   [Ohio](https://publicapps.odh.ohio.gov/EDW/DataBrowser/Browse/StateLayoutLockdownCancers)
    *   [Pennsylvania](https://www.phaim1.health.pa.gov/EDD/)

We also obtained age-adjusted incidence rates data for these three health
outcomes for the U.S. states New York, Ohio and Pennsylvania at the county-level
at fixed time points. The following bullet points provide links from 
where we obtained the data. The asthma data is provided for the year 2016 and
the skin cancer and lunger cancer data is average over the years 2014-2018.

*   Asthma
    *   Provide source
*   Melanoma of the skin (skin cancer) and lung cancer
    *   [All 50 states](https://statecancerprofiles.cancer.gov/quick-profiles/index.php?tabSelected=2&statename=newyork)

Each of these data sets were exported and each imported into R. For each
of the three health outcomes, the data sets were merged across the three states 
for the longitudinal data and for the cross-sectional data. This effort resulted in six 
data sets. The data import and cleaning of the original data can be found in the 
data import folder on our github repository.

We will now import the six data sets and merge all longitudinal data together
and merge all cross-sectional data together to have two data sets.

## Combining longitudinal state-level health outcome data

First, let's combine longitudinal lung and skin cancer data at the state level.
```{r warning=FALSE, message=FALSE}
# read in longitudinal state data for lung and skin cancer

# longitudinal state level lung cancer data
state_lc_data = read_csv("./data/lung/state_lc_data.csv")

# longitudinal state level melanoma data
state_mel_data = read_csv("./data/incidence_melanoma_data/state_melanoma_age_adjusted.csv") %>%
  rename(c("outcome" = "cancer_type"))

# combine cancer outcome data
lc_mel_state = bind_rows(state_lc_data, state_mel_data) %>%
  select(state, year, outcome, age_adjusted_incidence_rate)
```

Next, let's combine longitudinal asthma data with the longitudinal cancer data 
at the state level.

First, let's read in the longitudinal asthma state-level data.
```{r warning=FALSE, message=FALSE}
asthma_state <- read_csv(here::here("data", "asthma_state.csv")) %>%
  rename(age_adjusted_incidence_rate = rate) %>%
  select(state, year, outcome, age_adjusted_incidence_rate)
```

Next, let's combine the longitudinal asthma and longitudinal cancer state-level 
data.
```{r warning=FALSE}
# Bind rows of longitudinal asthma and longitudinal cancer by state and arrange 
# accordingly
lc_mel_asthma_state <- bind_rows(lc_mel_state, asthma_state) %>%
  arrange(state, year, outcome)
```

Lastly, let's write out the combined data.
```{r}
write_csv(lc_mel_asthma_state, here::here("data", "lc_mel_asthma_state.csv"))
```

## Combining cross-sectional county-level health outcome data

First, let's combine the cross-sectional lung and skin cancer data at the county 
level.
```{r warning=FALSE, message=FALSE}
# read in cross-sectional county data for lung and skin cancer

county_lc = read_csv("./data/lung/county_lc.csv")
county_mel = read_csv("./data/incidence_melanoma_data/state_county_melanoma_incidence_2014_2018.csv")

fips_codes = read_csv("./data/fips_codes.csv")

# cleaning to make the data sets compatible

county_mel = county_mel %>%
  rename(c("outcome" = "cancer_type")) %>%
  drop_na() 

county_mel <- county_mel[!(county_mel$county == "Ohio" | county_mel$county == "Pennsylvania" | 
     county_mel$county == "New York"), ] %>%
  filter(!grepl('SEER', county))

county_mel$county <- gsub(" County","", county_mel$county)

county_mel$county[county_mel$county == "St Lawrence"] <- "St. Lawrence"

county_lc$county <- gsub(" County","", county_lc$county)

# binding rows

lc_mel_county = bind_rows(county_lc, county_mel) %>%
  filter(complete.cases(.)) 


fips = fips_codes %>%
    janitor::clean_names() %>%
    rename(c("county" = "name"))

lc_mel_county = merge(lc_mel_county, fips, by = c("state", "county"))
```

Next, we will read in asthma data and combine it with the cancer data. 

First, let's read in the cross-sectional asthma county data, rename column 
names and add an additional column for county fips code.
```{r warning=FALSE, message=FALSE}
# Read in cross-sectional asthma data
# Only get county data for the year 2016 since we need county data at
# a fixed time point (i.e. not longitudinal)
# Adjust column names and types in prep for merge
asthma_county <- read_csv(here::here("data", "asthma_county.csv")) %>%
  filter(year == 2016) %>%
  select(-year) %>%
  rename(age_adjusted_incidence_rate = rate) %>%
  mutate(age_adjusted_incidence_rate = as.double(age_adjusted_incidence_rate))

asthma_county <- merge(asthma_county, fips, by = c("state", "county"))
```

Next, let's combine the cross-sectional asthma and cancer county data.
```{r warning=FALSE}
# Bind rows of cancer and asthma by county and arrange accordingly
lc_mel_asthma_county <- bind_rows(lc_mel_county, asthma_county) %>%
  arrange(state, county, outcome)
```

Lastly, let's write out the combined data.
```{r}
write_csv(lc_mel_asthma_county, here::here("data", "lc_mel_asthma_county.csv"))
```

Now, that we have merged the health outcomes data, we can explore.
```{r explore, include=FALSE}

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
counties <- rjson::fromJSON(file = url)


state_lc_wide = read_csv("./data/state_lc_wide.csv")

# plotly of county level data

fig <- plot_ly(state_lc_wide, x = ~year)
fig <- fig %>% add_lines(y = ~PA_AAIR, name = "Pennsylvania")
fig <- fig %>% add_lines(y = ~NY_AAIR, name = "New York")
fig <- fig %>% add_lines(y = ~OH_AAIR, name = "Ohio")
fig <- fig %>% layout(
    title = "Lung Cancer Age-Adjusted Incidence Rates",
    xaxis = list(
      rangeselector = list(
        buttons = list(
          list(
            count = 1,
            label = "1 yr",
            step = "year",
            stepmode = "backward"),
          list(
            count = 5,
            label = "5 yr",
            step = "year",
            stepmode = "backward"),
          list(
            count = 10,
            label = "10 yr",
            step = "year",
            stepmode = "backward"),
          list(step = "all"))),

      rangeslider = list(type = "year")),

    yaxis = list(title = "Age-Adjusted Incidence Rate (per 100,000"))

fig

lc_aqi_data = read_csv("./data/lc_aqi_data.csv")

fig1 <- lc_aqi_data %>%
  plot_ly(
    x = ~age_adjusted_incidence,
    y = ~y,
    frame = ~f,
    type = 'scatter',
    mode = 'markers',
    showlegend = F
  )

fig1 <- lc_aqi_data %>%
  plot_ly(
    x = ~avg_med_aqi, 
    y = ~age_adjusted_incidence_rate, 
    color = ~state, 
    frame = ~year, 
    text = ~state, 
    hoverinfo = "text",
    type = 'scatter',
    mode = 'markers'
  )
fig1 <- fig1 %>% layout(
  title = "Lung Cancer AAIR and Average Median AQI",
    xaxis = list(title = "Average Median AQI", 
      type = "log"
    ),
    yaxis = list(title = "Age Adjusted Incidence Rate")
  )


fig1

lc_fips = read_csv("./data/lung/lc_fips.csv")


county_map_lc <- plot_ly()
county_map_lc <- county_map_lc %>%
  add_trace(
    type = "choroplethmapbox",
    geojson = counties,
    locations = lc_fips$fips,
    z = lc_fips$age_adjusted_incidence_rate,
    text = lc_fips$county,
    colorscale = "Viridis",
    reversescale = TRUE,
    marker = list(line = list(
      width = 0),
      opacity = 0.5)
    )
county_map_lc <- county_map_lc %>%
  layout(
    title = "Age-Adjusted Lung Cancer Incidence Rates (2014-2018)",
    mapbox = list(
      style = "carto-positron",
      zoom = 4,
      center = list(lon = -77.215135, lat = 41.164818)
    )
  )


county_map_lc


state_melanoma_age_adjusted = read_csv("./data/incidence_melanoma_data/state_melanoma_age_adjusted.csv")

state_mel_plot = state_melanoma_age_adjusted %>%
  ggplot(aes(x = year, y = age_adjusted_incidence_rate, color = state)) +
  geom_line() +
  labs(
    title = "Age-adjusted incidence rate of Melanoma by Year colored by State",
    x = "Year",
    y = "Age-adjusted incidence rate of Melanoma", 
    color = "State"
  ) +
  scale_x_continuous(breaks = seq(1976, 2018, by = 4)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))



```



## Let's talk health outcomes:

To discuss how climate change has impacted the rate of certain negative health
outcomes, let's first assess their general prevalence.


# Recent County-Level Trends


Clearly, according to the map below, lung cancer is a relevant issue in 
our three states of interest (NY, PA, and OH). They also provide a good
range of examples, as we can see multiple counties with high age-adjusted 
incidence rates in Ohio, a few in New York, and virtually none in Pennsylvania.


```{r}

# plotly of county level data

county_map_lc

```


# Recent State Level Trends

We can also check how these rates have changed by state over time:



```{r}

# plotly of county level data

fig

state_mel_plot


```






