---
title: "merge_outcomes"
output: github_document
---

```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
```


## Combining state-level outcome data

First, let's combine lung and skin cancer data at the state level.
```{r}
# read in state data for lung and skin cancer

#state level lung cancer data
state_lc_data = read_csv("./data/lung/state_lc_data.csv")

#state level melanoma data
state_mel_data = read_csv("./data/incidence_melanoma_data/state_melanoma_age_adjusted.csv") %>%
  rename(c("outcome" = "cancer_type"))

# combine cancer outcome data
lc_mel_state = bind_rows(state_lc_data, state_mel_data)
```

Next, let's combine asthma data with the cancer data at the state level.
First, let's read in the asthma state-level data.
```{r}
asthma_state <- read_csv(here::here("data", "asthma_state.csv"))
```
Next, let's combine the asthma and cancer state-level data.
```{r}
# Bind rows of cancer and asthma by state and arrange accordingly
lc_mel_asthma_state <- bind_rows(lc_mel_state, asthma_state) %>%
  arrange(state, outcome)
```

Lastly, let's write out the combined data.
```{r}
write_csv(lc_mel_asthma_state, here::here("data", "lc_mel_asthma_state.csv"))
```

## Combining county-level outcome data

First, let's combine lung and skin cancer data at the county level.
```{r}
# read in county data for lung and skin cancer

county_lc = read_csv("./data/lung/county_lc.csv")
county_mel = read_csv("./data/incidence_melanoma_data/state_county_melanoma_incidence_2014_2018.csv")

fips_codes = read_csv("./data/fips_codes.csv")

# cleaning to make the data sets compatible

county_mel = county_mel %>%
  rename(c("outcome" = "cancer_type")) %>%
  drop_na() 

county_mel <- county_mel[!(county_mel$county == "Ohio" | county_mel$county == "Pennsylvania" | 
     county_mel$county == "New York"), ] %>%
  filter(!grepl('SEER', county))

county_mel$county <- gsub(" County","", county_mel$county)

county_mel$county[county_mel$county == "St Lawrence"] <- "St. Lawrence"

county_lc$county <- gsub(" County","", county_lc$county)

# binding rows

lc_mel_county = bind_rows(county_lc, county_mel) %>%
  filter(complete.cases(.)) 


fips = fips_codes %>%
    janitor::clean_names() %>%
    rename(c("county" = "name"))

lc_mel_county = merge(lc_mel_county, fips, by = c("state", "county"))
```

Next, we will read in asthma data and combine it with the cancer data. 

First, let's read in the asthma county data, rename column names and add an 
additional column for county fips code.
```{r}
# Read in asthma data
# Only get county data for the year 2016 since we need county data at
# a fixed time point (i.e. not longitudinal)
# Adjust column names and types in prep for merge
asthma_county <- read_csv(here::here("data", "asthma_county.csv")) %>%
  filter(year == 2016) %>%
  select(-year) %>%
  rename(age_adjusted_incidence_rate = rate) %>%
  mutate(age_adjusted_incidence_rate = as.double(age_adjusted_incidence_rate))

asthma_county <- merge(asthma_county, fips, by = c("state", "county"))
```
Next, let's combine the asthma and cancer county data.
```{r}
# Bind rows of cancer and asthma by county and arrange accordingly
lc_mel_asthma_county <- bind_rows(lc_mel_county, asthma_county) %>%
  arrange(state, county, outcome)
```

Lastly, let's write out the combined data.
```{r}
write_csv(lc_mel_asthma_county, here::here("data", "lc_mel_asthma_county.csv"))
```





