---
title: "Regression Analyses"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

#Part 1: State Level Trends, 2001-2015

## Data

We begin by reading in the combined health outcomes data and the combined air pollution and UV radiation data to produce an analysis at the state level. We quickly noticed that some of our variables in the air pollution and UV data were strongly correlated with each other, so we omitted several for the purposes of our regression analysis (collinearity violates our assumptions); the final correlation matrix is given below.

```{r}
library(tidyverse)

outcomes_state <- read_csv("data/lc_mel_asthma_state.csv")
outcomes_county <- read_csv("data/lc_mel_asthma_county.csv")
ap_uv <- read_csv("ap/ap_uv/apuv.csv")

#State-level analysis - fewer predictors
ap_uv_state_slim <- ap_uv %>%
  select(-c(countyfips, county,
            pm25_max_pred, pm25_mean_pred, pm25_pop_pred,
            o3_max_pred, o3_mean_pred, o3_pop_pred,
            i310, i305, i380, edr)) %>%
  group_by(state, year, season) %>%
  #Take medians across county for each season
  summarize(across(pm25_med_pred:edd, median)) %>%
  pivot_wider(names_from = season, values_from = pm25_med_pred:edd)

#State-level analysis - more predictors
ap_uv_state <- ap_uv %>%
  select(-c(countyfips, county)) %>%
  group_by(state, year, season) %>%
  #Take medians across county for each season
  summarize(across(pm25_max_pred:i380, median)) %>%
  pivot_wider(names_from = season, values_from = pm25_max_pred:i380)

outcomes_wider_state <- outcomes_state %>%
  pivot_wider(names_from = outcome, values_from = age_adjusted_incidence_rate)
  
state_analysis <- left_join(outcomes_wider_state, ap_uv_state_slim)

#Check for correlations
state_analysis_corr <- state_analysis %>% 
  select(`lung cancer`:edd_Winter) %>% 
  do(as.data.frame(cor(., method="pearson", use="pairwise.complete.obs"))) %>%
  knitr::kable()
```

## Lung Cancer

### Exploration

Before performing our analysis, we may want to look at some bivariate plots with trends over time to see if there are general trends from 2001 to 2015.

```{r}
library(GGally)

state_analysis %>%
  filter(year %in% 2001:2015) %>%
  ggpairs(columns = c("year","lung cancer", "o3_med_pred_Summer", "pm25_med_pred_Summer"),
          mapping = aes(group = state, color = state),
          columnLabels = c("Year", "Lung Cancer/100,000","Median summer O3, ppm", "PM 2.5, ug/m^3"))

  
state_plot <- state_analysis %>%
  ggplot(aes(x = year, y = `lung cancer`, group = state, color = state)) +
  geom_path()

spring_plot <- state_analysis %>%
  ggplot(aes(x = o3_med_pred_Spring, y = `lung cancer`, group = state, color = state)) +
  geom_point()

summer_plot <- state_analysis %>%
  ggplot(aes(x = o3_med_pred_Summer, y = `lung cancer`, group = state, color = state)) +
  geom_point()

fall_plot <- state_analysis %>%
  ggplot(aes(x = o3_med_pred_Fall, y = `lung cancer`, group = state, color = state)) +
  geom_point()

library(patchwork)
(state_plot + spring_plot) / (summer_plot + fall_plot)
```

### Model Fit

We proceed using backwards selection BIC from 2001 to 2015

```{r}
state_2001_2015 <- state_analysis %>%
  filter(year %in% 2001:2015) %>%
  #exclude asthma, lots of missingness
  select(-asthma)

fit <- lm(`lung cancer` ~ ., data = state_2001_2015)
step_bic <- step(fit, trace = 0, k = log(nobs(fit)), direction = "backward")
summary(step_bic)
```

Am I reading that right? $R^2$ of 0.98! That sounds too good to be true. Let's look at some diagnostic plots.

```{r}
lung_fit <- state_2001_2015 %>%
  modelr::add_residuals(step_bic) %>%
  modelr::add_predictions(step_bic)

lung_fit %>%
  ggplot(aes(x = pred, y = resid)) + geom_point() + labs(x = "Predicted value", y = "Residual")
```

Residuals appear to have a slight upward curvature overall. What variables are driving this?
```{r}
spring_resid <- lung_fit %>%
  ggplot(aes(x = o3_med_pred_Spring, y = resid)) + geom_point() + labs(x = "Spring Median O3",
                                                          y = "Residual")
summer_resid <- lung_fit %>%
  ggplot(aes(x = o3_med_pred_Summer, y = resid)) + geom_point() + labs(x = "Summer Median O3", y = "Residual")

fall_resid <- lung_fit %>%
  ggplot(aes(x = o3_med_pred_Fall, y = resid)) + geom_point() + labs(x = "Fall Median O3", y = "Residual")

state_resid <- lung_fit %>%
  ggplot(aes(x = state, y = resid)) + geom_violin() + labs(x = "State", y = "Residual")

(spring_resid + summer_resid) / (fall_resid + state_resid)
```

Hm...

```{r}
plot(step_bic, which = 2)
```

Normality assumption doesn't look so great.

```{r}
plot(step_bic, which = 5)
```

Datapoint number 28 is a borderline outlier.

### Cross validation

What happens if we cross-validate against simpler models? I suspect we are overfit. We will use leave-one-out cross-validation.

```{r}
library(modelr)
set.seed(15)
cv_df <- crossv_loo(state_2001_2015)
    
cv_df <- cv_df %>% 
  mutate(
    #Ozone only
    ozone  = map(train, ~lm(`lung cancer` ~ o3_med_pred_Spring + o3_med_pred_Summer +
                              o3_med_pred_Fall, data = .x)),
    #Ozone with year and state
    ozone_state_year  = map(train, ~lm(`lung cancer` ~ o3_med_pred_Spring + o3_med_pred_Summer +
                                         o3_med_pred_Fall + year + state, data = .x)),
    #Ozone with year and state with time interactions
    ozone_state_year2  = map(train, ~lm(`lung cancer` ~ year*(o3_med_pred_Spring +
                                                              o3_med_pred_Summer +
                                                              o3_med_pred_Fall) + year + state +
                                                              year*state, data = .x)),
    
    #BIC model
    bic_fit = map(train, ~lm(`lung cancer` ~ state + year + o3_med_pred_Fall +
                               o3_med_pred_Spring + o3_med_pred_Summer + edd_Fall + edd_Spring +
                               edd_Summer, data = .x))) %>% 
  mutate(
    rmse_ozone = map2_dbl(ozone, test, ~rmse(model = .x, data = .y)),
    rmse_ozone_state_year = map2_dbl(ozone_state_year, test, ~rmse(model = .x, data = .y)),
    rmse_ozone_state_year2 = map2_dbl(ozone_state_year2, test, ~rmse(model = .x, data = .y)),
    rmse_bic_fit = map2_dbl(bic_fit, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() + labs(x = "Model", y = "RMSE")
```

In retrospect, this model really isn't appropriate for these data - this is a time series. We should probably include time interaction effects, and maybe use generalized least squares. Also cross-validation with a rolling window may be better. But still, somehow including EDD—a UV radiation variable makes the model for lung cancer better for some reason??

