---
title: "UV radiation data"
author: "Jesse R. Ames"
date: "11/29/2021"
output: github_document
---

## Data

Source: https://data.cdc.gov/Environmental-Health-Toxicology/Population-Weighted-Ultraviolet-Irradiance-2004-20/h28b-t43q/data#About

Before downloading, I filtered to our states of interest (Maine, New York, Ohio, and Pennsylvania) using the CDC's web application. The state FIPS codes for Maine, New York, Ohio, and Pennsylvania are 23, 36, 39, and 42, respectively, per the U.S. Department of Agriculture.

```{r}
library(tidyverse)
uv <- read_csv("data/Population-Weighted_Ultraviolet_Irradiance__2004-2015.csv")
names(uv)
```

According to the metadata provided with this dataset (OMIUV_Metadata_May2017.pdf), we have the following variables:

* `statefips` - FIPS code of the state of the observations
* `countyfips` - FIPS code of the county of the observations
* `year`, `month`, and `day` (from 2004 to 2015)
* `edd` - "Daily county-level population weighted erythemally weighted daily dose, with the
unit of J/m^2"
* `edr` - "Daily county-level population weighted erythemally weighted irradiance at local
solar noon time, with the unit of mW/m^2"
* `i305` - "Daily county-level population weighted spectral irradiance at local solar noon time
at 305 nm, with the unit of mW/m^2/nm"
* `i310` - "Daily county-level population weighted spectral irradiance at local solar noon time
at 310 nm, with the unit of mW/m^2/nm"
* `i324` - "Daily county-level population weighted spectral irradiance at local solar noon time
at 324 nm, with the unit of mW/m^2/nm"
* `i380` - "Daily county-level population weighted spectral irradiance at local solar noon time
at 380 nm, with the unit of mW/m^2/nm"

The metadata additionally specifies how missing data was coded:

"NAs in the dataset appear as missing values; these refer to no valid data and are caused
by invalid retrievals from OMI. We suggest to treat NAs as missing."

Now, we need the county-level FIPS codes. How do we get those? Web scraping the [USDA website](https://www.nrcs.usda.gov/wps/portal/nrcs/detail/national/home/?cid=nrcs143_013697):

```{r}
library(rvest)
library(httr)

usda_countyfips <- read_html("https://www.nrcs.usda.gov/wps/portal/nrcs/detail/national/home/?cid=nrcs143_013697")

#Use SelectorGadget

countyfips_string <- usda_countyfips %>%
  html_elements(".data") %>%
  html_text2()

#Break this crazy string into useful data
countyfips_matrix <- matrix(data = unlist(strsplit(countyfips_string, split = "\r "))[-c(1:3)],
                            ncol = 3, byrow = TRUE)
countyfips_df <- tibble(countyfips = countyfips_matrix[,1],
                        county = countyfips_matrix[,2],
                        state = countyfips_matrix[,3]) %>%
  mutate(countyfips = as.double(countyfips))

#left_join to add county and state to the UV data
uv <- left_join(uv, countyfips_df, by = "countyfips")

#County, state
uv <- uv %>%
  mutate(county = paste(county, state, sep = ", "))

#Missing data?
skimr::skim(uv)

#What are these missing rows? Maybe it's the 2004 data?
missing <- uv %>%
  filter(is.na(edd))

missing %>% janitor::tabyl(year, month, state)

#We have several columns which are all highly correlated with one another
uv %>% 
  select(edd:i380) %>% 
  do(as.data.frame(cor(., method="pearson", use="pairwise.complete.obs")))
```

## Visualizations

```{r}
#Boxplots
uv %>%
  ggplot(aes(x = state, y = edd, color = state)) +
  geom_boxplot() +
  labs(title = "Comparison of Erythemal Daily Dose by State",
       x = "State", y = "EDD (J/m^2)")

#Density
uv %>%
  ggplot(aes(x = edd, color = state)) +
  geom_density() +
  labs(title = "Distribution of Erytemal Daily Dose by State",
       x = "EDD (J/m^2)")

#County breakdown
uv %>%
  filter(state == "NY") %>%
  ggplot(aes(x = reorder(county,edd, na.rm = TRUE), y = edd)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(title = "Comparison of Erythemal Daily Dose by County, NY",
       x = "County", y = "EDD (J/m^2)")

#Monthly trends
uv %>%
  mutate(month = as.character(month)) %>%
  mutate(month = fct_recode(month,
                            Jan = "1", Feb = "2", Mar = "3", Apr = "4",
                            May = "5", Jun = "6", Jul = "7", Aug = "8",
                            Sep = "9", Oct = "10", Nov = "11", Dec = "12")) %>%
  mutate(month = factor(month, levels = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug",
                                          "Sep","Oct","Nov","Dec"))) %>%
  ggplot(aes(x = month, y = edd, color = state)) +
  geom_boxplot() +
  labs(title = "Comparison of Erythemal Daily Dose by State by Month")

#Annual trends
uv %>%
  mutate(year = as.factor(year)) %>%
  ggplot(aes(x = year, y = edd, color = state)) +
  geom_boxplot() +
  labs(title = "Comparison of Erythemal Daily Dose by State by Year")
```

Note that the data from 2004 start in October; they aren't full-year data.



